# RProject

# Part 1

## Prep workspace
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("ggplot2")) install.packages("ggplot2")
library(ggplot2)
if (!require("gtools")) install.packages("gtools")
library(gtools)
```

##Read in files
```{r}
SNP <- read_tsv("https://raw.githubusercontent.com/EEOB-BioData/BCB546X-Fall2019/master/assignments/UNIX_Assignment/snp_position.txt")
Fang <- read_tsv("https://raw.githubusercontent.com/EEOB-BioData/BCB546X-Fall2019/master/assignments/UNIX_Assignment/fang_et_al_genotypes.txt")
```

##View files
```{r}
view(SNP)
view(Fang)
```

#Prep directories for each species
```{r}
dir.create("Maize")
dir.create("Teosinte")
dir.create("graphs")
```

##Keep only columns 1, 3, and 4 from SNP data
```{r}
snippedSNP <- SNP[c(1,3,4)]
```

##Create separate files for Maize and Teosinte based on group name, and remove extra columns
```{r}
Teo <- filter(Fang, Group == "ZMPJA" | Group == "ZMPIL" | Group == "ZMPBA")
Maize <- filter(Fang, Group == "ZMMMR" | Group == "ZMMLR" | Group == "ZMMIL")
cut_Teo <- Teo[,-(1:3)]
cut_Maize <- Maize[,-(1:3)]
```

##Transposed Teosinte and Maize files. Add new columns for SNP ID
```{r}
tTeo <- as.data.frame(t(cut_Teo))
tMaize <- as.data.frame(t(cut_Maize))
tTeo <- rownames_to_column(tTeo,"SNP_ID")
tMaize <- rownames_to_column(tMaize,"SNP_ID")
```

##Merged SNP file with the Teosinte and Maize files
```{r}
mTeo <- merge(snippedSNP,tTeo, by.x = "SNP_ID", by.y = "SNP_ID")
mMaize <- merge(snippedSNP, tMaize, by.x = "SNP_ID", by.y = "SNP_ID")
```

##Ordered files by chromosome file, with ties broken by position number. There seem to have been problems when sorting by position
```{r}
mTeo <- mTeo[order(mTeo$Chromosome, mTeo$Position, decreasing = FALSE),]
```

##Creates new files for each Teosinte Chromosome. I wanted to write a for loop, but couldn't get it right.
```{r}
filter(mTeo, Chromosome == 1) %>% write_csv(path = "Teosinte/TEO_CHR1_inc.csv")
filter(mTeo, Chromosome == 2) %>% write_csv(path = "Teosinte/TEO_CHR2_inc.csv")
filter(mTeo, Chromosome == 3) %>% write_csv(path = "Teosinte/TEO_CHR3_inc.csv")
filter(mTeo, Chromosome == 4) %>% write_csv(path = "Teosinte/TEO_CHR4_inc.csv")
filter(mTeo, Chromosome == 5) %>% write_csv(path = "Teosinte/TEO_CHR5_inc.csv")
filter(mTeo, Chromosome == 6) %>% write_csv(path = "Teosinte/TEO_CHR6_inc.csv")
filter(mTeo, Chromosome == 7) %>% write_csv(path = "Teosinte/TEO_CHR7_inc.csv")
filter(mTeo, Chromosome == 8) %>% write_csv(path = "Teosinte/TEO_CHR8_inc.csv")
filter(mTeo, Chromosome == 9) %>% write_csv(path = "Teosinte/TEO_CHR9_inc.csv")
filter(mTeo, Chromosome == 10) %>% write_csv(path = "Teosinte/TEO_CHR10_inc.csv")
```

#Repeat sorting process for Maize
```{r}
mMaize <- mMaize[order(mMaize$Chromosome, mMaize$Position, decreasing = F),]
```

##Repeat file splicing for Maize
```{r}
filter(mMaize, Chromosome == 1) %>% write_csv(path = "Maize/MAIZE_CHR1_inc.csv")
filter(mMaize, Chromosome == 2) %>% write_csv(path = "Maize/MAIZE_CHR2_inc.csv")
filter(mMaize, Chromosome == 3) %>% write_csv(path = "Maize/MAIZE_CHR3_inc.csv")
filter(mMaize, Chromosome == 4) %>% write_csv(path = "Maize/MAIZE_CHR4_inc.csv")
filter(mMaize, Chromosome == 5) %>% write_csv(path = "Maize/MAIZE_CHR5_inc.csv")
filter(mMaize, Chromosome == 6) %>% write_csv(path = "Maize/MAIZE_CHR6_inc.csv")
filter(mMaize, Chromosome == 7) %>% write_csv(path = "Maize/MAIZE_CHR7_inc.csv")
filter(mMaize, Chromosome == 8) %>% write_csv(path = "Maize/MAIZE_CHR8_inc.csv")
filter(mMaize, Chromosome == 9) %>% write_csv(path = "Maize/MAIZE_CHR9_inc.csv")
filter(mMaize, Chromosome == 10) %>% write_csv(path = "Maize/MAIZE_CHR10_inc.csv")
```

##Sorted maize in decreasing order. I found the find and replace code on Slack; I don't know how it works.
```{r}
mMaize <- mMaize[order(mMaize$Chromosome,mMaize$Position, decreasing = T),]
mMaize <- apply(X = mMaize, MARGIN = 2, FUN = as.character)
mMaize[mMaize == "?/?"] <- "-/-"
mMaize <- as.data.frame(mMaize)
```

##repeat file splicing for descended sorting
```{r}
filter(mMaize, Chromosome == 1) %>% write_csv(path = "Maize/MAIZE_CHR1_dec.csv")
filter(mMaize, Chromosome == 2) %>% write_csv(path = "Maize/MAIZE_CHR2_dec.csv")
filter(mMaize, Chromosome == 3) %>% write_csv(path = "Maize/MAIZE_CHR3_dec.csv")
filter(mMaize, Chromosome == 4) %>% write_csv(path = "Maize/MAIZE_CHR4_dec.csv")
filter(mMaize, Chromosome == 5) %>% write_csv(path = "Maize/MAIZE_CHR5_dec.csv")
filter(mMaize, Chromosome == 6) %>% write_csv(path = "Maize/MAIZE_CHR6_dec.csv")
filter(mMaize, Chromosome == 7) %>% write_csv(path = "Maize/MAIZE_CHR7_dec.csv")
filter(mMaize, Chromosome == 8) %>% write_csv(path = "Maize/MAIZE_CHR8_dec.csv")
filter(mMaize, Chromosome == 9) %>% write_csv(path = "Maize/MAIZE_CHR9_dec.csv")
filter(mMaize, Chromosome == 10) %>% write_csv(path = "Maize/MAIZE_CHR10_inc.csv")
```

## Repeat resorting for teosinte and replace characters
```{r}
mTeo <- mTeo[order(mTeo$Chromosome,mTeo$Position, decreasing = T),]
mTeo <- apply(X = mTeo, MARGIN = 2, FUN = as.character)
mTeo[mTeo == "?/?"] <- "-/-"
mTeo <- as.data.frame(mTeo)
```

##Repeat splicing for descending teosinte
```{r}
filter(mTeo, Chromosome == 1) %>% write_csv(path = "Teosinte/TEO_CHR1_dec.csv")
filter(mTeo, Chromosome == 2) %>% write_csv(path = "Teosinte/TEO_CHR2_dec.csv")
filter(mTeo, Chromosome == 3) %>% write_csv(path = "Teosinte/TEO_CHR3_dec.csv")
filter(mTeo, Chromosome == 4) %>% write_csv(path = "Teosinte/TEO_CHR4_dec.csv")
filter(mTeo, Chromosome == 5) %>% write_csv(path = "Teosinte/TEO_CHR5_dec.csv")
filter(mTeo, Chromosome == 6) %>% write_csv(path = "Teosinte/TEO_CHR6_dec.csv")
filter(mTeo, Chromosome == 7) %>% write_csv(path = "Teosinte/TEO_CHR7_dec.csv")
filter(mTeo, Chromosome == 8) %>% write_csv(path = "Teosinte/TEO_CHR8_dec.csv")
filter(mTeo, Chromosome == 9) %>% write_csv(path = "Teosinte/TEO_CHR9_dec.csv")
filter(mTeo, Chromosome == 10) %>% write_csv(path = "Teosinte/TEO_CHR10_inc.csv")
```

#Part 2

##Add reshape2 to the environment
```{r}
library(reshape2)
```

##Melt maize and teosinte files. I don't know what this does.
```{r}
melt(mMaize, na.rm = T,)
melt(mTeo, na.rm = T,)
```

##Creates plots for maize and teosinte data based on SNP count per chromosome. Both plots look the same. I think I did something wrong. Saves images to "Graphs" directory
```{r}
ggplot(data = mMaize) + geom_bar(mapping = aes(x = mMaize$Chromosome)) + xlab(label = "Chromosome") + ylab(label = "Number of SNPs")
ggsave(filename = "graphs/MaizeSNPs.png")
ggplot(data = mTeo) + geom_bar(mapping = aes(x = mTeo$Chromosome)) + xlab(label = "Chromosome") +ylab(label = "Number of SNPs")
ggsave(filename = "graphs/TeosinteSNPs.png")
```

###I couldn't figure out the rest of the plotting
